Coherent optical creation of a single molecule

### 2020-12-15 paper.tex
Update on Overleaf.

\newcommand{\harvardphysics}{\affiliation{Department of Physics, Harvard University, Cambridge, Massachusetts 02138, USA}}
\newcommand{\harvardccb}{\affiliation{Department of Chemistry and Chemical Biology, Harvard University, Cambridge, Massachusetts 02138, USA}}
\newcommand{\cua}{\affiliation{Harvard-MIT Center for Ultracold Atoms, Cambridge, Massachusetts 02138, USA}}
\newcommand{\agendile}{\affiliation{Agendile LLC, Cambridge, Massachusetts 02139, USA}}
\newcommand{\gradstudent}{
  \harvardphysics
  \harvardccb
  \cua
}

@@

\begin{document}
% \title{Coherent optical association of atoms into a single molecule}
\title{Coherent optical creation of a single molecule}
\author{Yichao~Yu}
\thanks{Y.Y.~and K.W.~contributed equally to this work.}
\email{yichaoyu@g.harvard.edu}
\gradstudent
\author{Kenneth~Wang}
\thanks{Y.Y.~and K.W.~contributed equally to this work.}
\gradstudent
\author{Jonathan~D.~Hood}
\affiliation{Department of Chemistry, Purdue University, West Lafayette, Indianna, 47906, USA}
\author{Lewis~R.~B.~Picard}
\gradstudent
\author{Jessie~T.~Zhang}
\gradstudent
\author{William~B.~Cairncross}
\harvardccb
\harvardphysics
\cua
\author{Jeremy~M.~Hutson}
\affiliation{Joint Quantum Centre Durham-Newcastle, Department of Chemistry, Durham University, Durham, DH1 3LE, United Kingdom}
\author{Rosario Gonzalez-Ferez}
\affiliation{Instituto Carlos I de F\'{\i}sica Te\'orica y Computacional, and Departamento de F\'{\i}sica At\'omica, Molecular y Nuclear,  Universidad de Granada, 18071 Granada, Spain}
\author{Till Rosenband}
\agendile
\author{Kang-Kuen~Ni}
\email{ni@chemistry.harvard.edu}
\harvardccb
\harvardphysics
\cua

@@

Here, we demonstrate coherent association of an atom pair to a weakly bound molecule
by two-photon optical Raman transfer via an electronic excited state,
as shown in Fig.~\ref{f-theory}a.
The technique does not rely on a Feshbach resonance,
molecular states bound by a only few~MHz, or a narrow excited state.
The resulting single molecule is in a well-defined internal quantum state
and predominantly in its motional ground state.
A vibrational state of the electronic excited state $\mathrm{c^3\Sigma^+}(\Omega = 1)$
serves as the intermediate state in the Raman technique,
and is chosen to minimize photon scattering during Raman Rabi oscillations.
To reduce photon scattering and sensitivity to laser intensity noise further,
we choose the initial and final states that balance the two Rabi frequencies as much as possible.
This approach applies to a variety of molecules that can be created atom-by-atom
with full quantum-state control.

@@

We optimize over intermediate states by calculating the Raman Rabi frequency $\Omega_{\rm R}$
and scattering rate $\Gamma_{\rm s}$ at different detunings from the atomic threshold,
taking into account of all states of
8 excited molecular potentials~\cite{Korek2007, Grochola2011, Zaharova2009, Grochola2010, Zabawa2012}
and the continuum~\cite{Liu2017}.
This calculation shows that the figure of merit $\Omega_{\rm R}/\Gamma_{\rm s}$
can be larger for the lowest vibrational state compared to higher bound states
at a cost of a smaller transfer rate $\Omega_{\rm R}$, as shown in Fig.~\ref{f-theory}c;
see Supplementary Material for details.
As a result, we choose the $v'=0$ level of $\mathrm{c^3\Sigma^+}(\Omega = 1)$
as the intermediate state near which to drive Raman transitions.

@@

\begin{figure}[t!]
  \includegraphics[width=0.48\textwidth]{imgs/fig-raman.pdf}
  \caption{Coherent transfer of atoms to molecules.  The molecular state is dark to the imaging step and corresponds to zero signal.
    (a) Raman difference frequency scans for various durations
    showing the resonance as a function of the detuning from
    the fitted resonance at $770.5715(1)$~MHz, near the calculated value of $763$~MHz.
    (b) Raman pulse-length scan on resonance.
    A decaying Rabi oscillation shows the coherence of
    the Raman transfer process.
    A model is fitted to (a) and (b) to determine
    the Raman Rabi frequency and loss rates.
    \emph{Inset:} Geometry and polarization of trap and Raman beam relative to the magnetic field.
    The 3.25~mW beam is focused to a waist of $0.9~\mathrm{\mu m}$
    that confines the atoms and molecule.
    A perpendicular $B_0=8.83(2)~\mathrm{G}$ magnetic field
    defines the quantization axis and the atoms experience predominantly $\pi$-polarized light.
    \label{f-raman}}
\end{figure}
After merging the tweezers, we drive the atoms into spin combination $\ket{\uparrow_{\Na} \downarrow_{\Cs}}$ with a large scattering length
by performing a Cs spin flip while taking into account
the $-30.7~\mathrm{kHz}$ interaction shift~\cite{Hood2019}.
This is the initial atomic state for Raman transfer.
The spin flip selectively transfers atoms in the relative motional ground state,
removing any background from atoms in excited states of relative motion
\footnote{This interaction shift is larger than the differential axial trapping frequency
  between Na and Cs atoms, which decouples the relative and center of mass motional state
  and improves the robustness of our preparation of the relative motional ground state.}.
For the experiment reported here,
$31\mathrm{\%}$ of initial two-atom population is transferred. Of this population, over $60\%$ is in the ground state of center-of-mass motion, inferred from Raman sideband thermometry.

@@

@@

The molecular lifetime measurement in Fig.~\ref{f-lifetime}a is performed
by preparing the molecule with a $\pi$-pulse,
followed after a variable delay by a second dissociating $\pi$-pulse.
Together with the Rabi oscillation in Fig.~\ref{f-raman}b,
we obtain a molecule lifetime of $0.199(9)~\mathrm{ms}$.
Preliminary experiments and theoretical considerations indicate that
the molecular lifetime may be limited by two-photon coupling
to the atomic continuum~\cite{YichaoYu}.
Atom loss is shown to be small in Fig.~\ref{f-lifetime}b
by measuring the two-body lifetimes of the atoms directly
without the second Raman frequency.
In principle, destructive interference that reduces the Raman Rabi rate $\Omega_{\rm R}$
for negative detunings $\Delta$ changes to constructive interference
for positive detunings, but additional molecular resonances make the positive region unusable.
More negative detunings that might reduce the scattering rate
were prevented by vanishing of $\Omega_{\rm R}$.

@@

In conclusion, we have coherently formed a weakly bound NaCs molecule in an optical tweezer
by optical Raman transfer.  This process is enabled by utilizing a deeply-bound intermediate state, as well as highly-interacting initial atomic states.  A theoretical investigation including 8 excited state potentials,
the excited atomic continuum, and coupled-channel ground-state wavefunctions indicates
the potential for higher transfer efficiency than the observed value of $69$~\%.  Future experiments may benefit from better balancing of the up-leg and down-leg Rabi frequencies, for example by driving to more deeply bound states.  If possible, destructive interference that reduces the two-photon Rabi rate should be avoided.  Nonlinear optical effects that limit the molecular state lifetime can also be explored.

@@

\begin{acknowledgments}
  We thank Bo Gao and Paul Julienne for discussion and Robert Moszynski for providing theoretical transition dipole moments of NaCs. This work is supported by the NSF~(PHY-1806595), the AFOSR~(FA9550-19-1-0089), ARO DURIP (W911NF1810194) and the Arnold and Mabel Beckman foundation. J.~T.~Z. is supported by a National Defense Science and Engineering Graduate Fellowship. W.~C. is supported by a Max Planck-Harvard Research Center for Quantum Optics fellowship. K.~W. is supported by an NSF GRFP fellowship. J.~M.~H. is supported by the U.K. Engineering and Physical Sciences Research Council (EPSRC) Grants No.\ EP/N007085/1, EP/P008275/1 and EP/P01058X/1.
\end{acknowledgments}

\TR{What is meant by ``result is consistent with the decay of the Rabi oscillation'' if independent fits not done?  See also Fig.~5 caption.}

>>>>

\newcommand{\harvardphysics}{\affiliation{Department of Physics, Harvard University, Cambridge, Massachusetts 02138, USA}}
\newcommand{\harvardccb}{\affiliation{Department of Chemistry and Chemical Biology, Harvard University, Cambridge, Massachusetts 02138, USA}}
\newcommand{\cua}{\affiliation{Harvard-MIT Center for Ultracold Atoms, Cambridge, Massachusetts 02138, USA}}
\newcommand{\agendile}{\affiliation{Agendile LLC, Cambridge, Massachusetts 02139, USA}}
\newcommand{\itamp}{\affiliation{ITAMP, Harvard-Smithsonian Center for Astrophysics, Cambridge, Massachusetts 02138, USA}}
\newcommand{\gradstudent}{
  \harvardphysics
  \harvardccb
  \cua
}

@@

\begin{document}
% \title{Coherent optical association of atoms into a single molecule}
\title{Coherent optical creation of a single molecule}
\author{Yichao~Yu}
\thanks{Y.Y.~and K.W.~contributed equally to this work.}
\gradstudent
\author{Kenneth~Wang}
\thanks{Y.Y.~and K.W.~contributed equally to this work.}
\gradstudent
\author{Jonathan~D.~Hood}
\affiliation{Department of Chemistry, Purdue University, West Lafayette, Indianna, 47906, USA}
\author{Lewis~R.~B.~Picard}
\gradstudent
\author{Jessie~T.~Zhang}
\gradstudent
\author{William~B.~Cairncross}
\harvardccb
\harvardphysics
\cua
\author{Jeremy~M.~Hutson}
\affiliation{Joint Quantum Centre Durham-Newcastle, Department of Chemistry, Durham University, Durham, DH1 3LE, United Kingdom}
\author{Rosario Gonzalez-Ferez}
\affiliation{Instituto Carlos I de F\'{\i}sica Te\'orica y Computacional, and Departamento de F\'{\i}sica At\'omica, Molecular y Nuclear,  Universidad de Granada, 18071 Granada, Spain}
\itamp
\author{Till Rosenband}
\agendile
\author{Kang-Kuen~Ni}
\email[To whom correspondence should be addressed: ]{ni@chemistry.harvard.edu}
\harvardccb
\harvardphysics
\cua

@@

Here, we demonstrate coherent association of an atom pair to a weakly bound molecule
by two-photon optical Raman transfer via an electronic excited state,
as shown in Fig.~\ref{f-theory}a, taking NaCs as the prototype system.
The technique does not rely on a Feshbach resonance,
molecular states bound by a only few~MHz, or a narrow excited state.
The resulting single molecule is in a well-defined internal quantum state
and predominantly in its motional ground state.
A vibrational state of the electronic excited state $\mathrm{c^3\Sigma^+}(\Omega = 1)$
serves as the intermediate state in the Raman technique,
and is chosen to minimize photon scattering during Raman Rabi oscillations.
To reduce photon scattering and sensitivity to laser intensity noise further,
we choose the initial and final states that balance the two Rabi frequencies as much as possible.
This approach applies to a variety of molecules that can be created atom-by-atom
with full quantum-state control.

@@

We optimize over intermediate states by calculating the total Raman Rabi frequency $\Omega_{\rm R}$
and scattering rate $\Gamma_{\rm s}$ at different detunings from the atomic threshold,
taking into account all states of
8 excited molecular electronic potentials~\cite{Korek2007, Grochola2011, Zaharova2009, Grochola2010, Zabawa2012}
and the continuum~\cite{Liu2017}; 
see Supplementary Material for details.
This calculation shows that the figure of merit $\Omega_{\rm R}/\Gamma_{\rm s}$
can be larger for the lowest vibrational state compared to higher bound states
at a cost of a smaller transfer rate $\Omega_{\rm R}$, as shown in Fig.~\ref{f-theory}c.
As a result, we choose the $v'=0$ level of $\mathrm{c^3\Sigma^+}(\Omega = 1)$
as the intermediate state near which to drive Raman transitions.

@@

\begin{figure}[t!]
  \includegraphics[width=0.48\textwidth]{imgs/fig-raman.pdf}
  \caption{Coherent transfer of atoms to molecules.  The molecular state is dark to the imaging step and corresponds to zero signal.
    (a) Raman difference frequency scans for various durations
    showing the resonance as a function of the detuning from
    the fitted resonance at $770.5715(1)$~MHz, near the calculated value of $763$~MHz.
    (b) Raman pulse-length scan on resonance.
    A decaying Rabi oscillation shows the coherence of
    the Raman transfer process.
    A model is fitted to (a) and (b) to determine
    the Raman Rabi frequency and loss rates.
    \emph{Inset:} Geometry and polarization of trap and Raman beam relative to the magnetic field.
    The 3.25~mW beam is focused to a waist of $0.9~\mathrm{\mu m}$
    that confines the atoms and molecule.
    A perpendicular $B_0=8.83(2)~\mathrm{G}$ magnetic field
    defines the quantization axis and the atoms experience predominantly $\pi$-polarized light.
    \label{f-raman}}
\end{figure}

@@

After merging the tweezers, we drive the atoms into spin combination $\ket{\uparrow_{\Na} \downarrow_{\Cs}}$ with a large scattering length
by performing a Cs spin flip while taking into account
the $-30.7~\mathrm{kHz}$ interaction shift~\cite{Hood2019}.
This is the initial atomic state for Raman transfer.
The spin flip selectively transfers atoms in the relative motional ground state,
removing any background from atoms in excited states of relative motion
\footnote{This interaction shift is larger than the differential axial trapping frequency
  between Na and Cs atoms, which decouples the relative and center of mass motional state
  and improves the robustness of our preparation of the relative motional ground state.}.
For the experiment reported here,
$31\mathrm{\%}$ of initial two-atom population is transferred. Of this population, over $60\%$ is in the ground state of center-of-mass motion, inferred from Raman sideband thermometry.

@@

The molecular lifetime measurement in Fig.~\ref{f-lifetime}a is performed
by preparing the molecule with a $\pi$-pulse,
followed after a variable delay by a second dissociating $\pi$-pulse. The measured lifetime of $0.196(14)~\mathrm{ms}$ is consistent with the decaying time scale of  $0.203(13)~\mathrm{ms}$ of the  Rabi oscillation 
 in Fig.~\ref{f-raman}b. %,we obtain a molecule lifetime of $0.199(9)~\mathrm{ms}$.
Preliminary experiments and theoretical considerations indicate that
the molecular lifetime may be limited by two-photon coupling
to the atomic continuum~\cite{YichaoYu}.
Atom loss is shown to be small in Fig.~\ref{f-lifetime}b
by measuring the two-body lifetimes of the atoms directly
without the second Raman frequency.
In principle, destructive interference that reduces the Raman Rabi rate $\Omega_{\rm R}$
for negative detunings $\Delta$ changes to constructive interference
for positive detunings, but additional molecular resonances make the positive region unusable.
More negative detunings that might reduce the scattering rate
were prevented by vanishing of $\Omega_{\rm R}$.

@@

In conclusion, we have coherently formed a weakly bound NaCs molecule in an optical tweezer
by optical Raman transfer.  This process is enabled by utilizing a deeply-bound intermediate state, as well as highly-interacting initial atomic states.  A theoretical investigation including 8 excited molecular electronic potentials,
the electronically excited atomic continuum, and coupled-channel ground-state wavefunctions indicates
the potential for higher transfer efficiency than the observed value of $69$~\%.  Future experiments may benefit from better balancing of the up-leg and down-leg Rabi frequencies, for example by driving to more deeply bound states.  If possible, destructive interference that reduces the two-photon Rabi rate should be avoided.  Nonlinear optical effects that limit the molecular state lifetime can also be explored.

@@

\begin{acknowledgments}
  We thank Bo Gao and Paul Julienne for discussion and Robert Moszynski for providing theoretical transition dipole moments of NaCs. This work is supported by the NSF~(PHY-1806595), the AFOSR~(FA9550-19-1-0089), ARO DURIP (W911NF1810194) and the Arnold and Mabel Beckman foundation. J.~T.~Z. is supported by a National Defense Science and Engineering Graduate Fellowship. W.~C. is supported by a Max Planck-Harvard Research Center for Quantum Optics fellowship. K.~W. is supported by an NSF GRFP fellowship. J.~M.~H. is supported by the U.K. Engineering and Physical Sciences Research Council (EPSRC) Grants No.\ EP/N007085/1, EP/P008275/1 and EP/P01058X/1. R.~G.~F. acknowledges financial support of the Spanish Project FIS2017-89349-P (MINECO), and the Andalusian research group FQM-207.
\end{acknowledgments}

%\TR{What is meant by ``result is consistent with the decay of the Rabi oscillation'' if independent fits not done?  See also Fig.~5 caption.}
